---
title: "t tests"
output: html_notebook
date: "2025-06-05"
---

do paired t-tasts on the gradient metrics

```{r}
library(afex)
library(lme4)
library(lmerTest)
library(readxl)
work_dir <- "/Users/apurplespoon/Desktop/UCAS/YanLab/Suzhou_Rumination/Revise/GradMetrix"
df <- read_excel(paste(work_dir, "/RumBeijing.xlsx", sep = ""))
df$sub_idx <- factor(df$sub_idx)
df$group <- factor(df$group)
df$condition <- factor(df$condition)
```

using LME

```{r}
df$condition <- relevel(df$condition, ref = "2")
model <- lmer(Dispersion ~ condition + head_motion + (1|sub_idx),data = df)

# 拟合 full model（含 condition）
full_model <- lmer(Dispersion ~ condition + head_motion + (1 | sub_idx),
                   data = df, REML = FALSE)

# 拟合 reduced model（去掉 condition）
reduced_model <- lmer(Dispersion ~ head_motion + (1 | sub_idx),
                      data = df, REML = FALSE)

# 残差平方和
rss_full <- sum(residuals(full_model)^2)
rss_reduced <- sum(residuals(reduced_model)^2)

# pseudo R²（近似固定效应解释率）
r2 <- (rss_reduced - rss_full) / rss_reduced

# 计算 Cohen's f²
f2 <- r2 / (1 - r2)

# 输出结果
summary(model)
cat("✅ Cohen’s f² for condition is:", round(f2, 4), "\n")

```

do two sample t-tasts on the gradient metrics
```{r}
library(afex)
library(lme4)
library(lmerTest)
library(readxl)
work_dir <- "/Users/apurplespoon/Desktop/UCAS/YanLab/Suzhou_Rumination/Revise/GradMetrix"
df <- read_excel(paste(work_dir, "/RUMTwottest.xlsx", sep = ""))
df$sub_idx <- factor(df$sub_idx)
df$group <- factor(df$group)
df$condition <- factor(df$condition)
```

using lm()
```{r}
df$group <- factor(df$group)
df$group <- relevel(df$group, ref = "2")
model <- lm(Dispersion ~ group + head_motion, data = df)

# 完整模型（包含 group + head_motion）
full_model <- lm(Dispersion ~ group + head_motion, data = df)

# 简化模型（只包含 head_motion）
reduced_model <- lm(Dispersion ~ -head_motion, data = df)

# 提取残差平方和
rss_full <- sum(residuals(full_model)^2)
rss_reduced <- sum(residuals(reduced_model)^2)

# 计算 pseudo R² 和 f²
r2 <- (rss_reduced - rss_full) / rss_reduced
f2 <- r2 / (1 - r2)

# 输出结果

summary(model)
cat("✅ Cohen's f² for group is:", round(f2, 4), "\n")

```

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
